Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a  Mesoscale Convective System Case During RELAMPAGO campaign

### 2022-05-03 analysis/chapters/paper_ana_20181122.Rmd
Revised version

---
title: "Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a  Mesoscale Convective System Case During RELAMPAGO"
author:
- name: Paola Belén Corrales
  email: paola.corrales@cima.fcen.uba.ar
  affiliation: UBA,CIMA,CNRS
  footnote: 1
- name: Victoria Galligani
  affiliation: UBA,CIMA,CNRS
- name: Juan Ruiz
  affiliation: UBA,CIMA,CNRS
- name: Luiz Sapucci
  affiliation: INPE
- name: María Eugenia Dillon
  affiliation: SMN,CONICET
- name: Yanina García Skabar
  affiliation: SMN,CONICET,CNRS
- name: Maximiliano Sacco
  affiliation: SMN
- name: Craig S. Schwartz
  affiliation: NCAR
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_book: 
      base_format: rticles::elsevier_article
abstract: |
  In this paper, we evaluate the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF over central and north-eastern Argentina where the surface and upper air observing networks are relatively coarse. We conducted a case study corresponding to a huge mesoscale convective system (MCS) developed during November 22, 2018. The accumulated precipitation associated with this MCS was quite high, exceeding 200 mm over northern Argentina and Paraguay. The MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign during November 2018.

  We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-members multiphisics  ensemble. We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files, AWS combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds and RAD, including SATWND and satellite radiances from different microwave and infrared sensors (AMSU, HIRS, MHS, ATMS, AIRS, and IASI). We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study. The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation. We also run ensemble forecasts initialized from each experiment show and evaluated their skill to predict precipitation. We found that the assimilated observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.  
journal: Atmospheric Research
#layout: authoryear,preprint,review,12pt
layout: final,5p,times,twocolumn,authoryear
address:
- code: UBA
  address: Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmósfera y los Océanos. Buenos Aires, Argentina.
- code: CIMA
  address: CONICET – Universidad de Buenos Aires. Centro de Investigaciones del Mar y la Atmósfera (CIMA). Buenos Aires, Argentina.
- code: CNRS
  address: CNRS – IRD – CONICET – UBA. Instituto Franco-Argentino para el Estudio del Clima y sus Impactos (IRL 3351 IFAECI). Buenos Aires, Argentina.
- code: SMN
  address: Servicio Meteorológico Nacional de Argentina.
- code: CONICET
  address: CONICET (Consejo Nacional de Investigaciones Científicas y Técnicas).
- code: INPE
  address: National Institute for Space Research, Brazil, Center for Weather Forecasting and Climate Studies.
- code: NCAR
  address: National Center for Atmospheric Research, Boulder, Colorado.
bibliography: 
  - paper_ANA.bib 
  - packages.bib
csl: elsevier-harvard.csl
footnote:
- code: 1
  text: Corresponding Author
keywords: Regional Data Assimilation, Surface Obsevations, Satellite Observations 
numbersections: true
# header-includes: 
#  - \usepackage{gensymb}
#  - \usepackage{subfig}
#  - \usepackage[inline]{showlabels}
#  - \usepackage{chngcntr}
#  - \usepackage{natbib}
#  - \usepackage{lineno}
#  - \linenumbers 
---

@@

Severe weather events cause significant human and material losses around the world. A large number of these phenomena are associated with the occurrence of deep moist convection including tornadoes, intense wind gusts, extreme precipitation in short time periods, large hail, and lightning. 
Southern South America has been shown to have one of the highest frequencies in the world of favorable conditions for high-impact meteorological events [@brooks2003] and large hail events [@cecil2012], particularly during austral spring and summer. 
This is also confirmed by observational evidence and high impact weather reports [@matsudo2015; @rasmussen2014]. Recently, the RELAMPAGO-CACTI field campaign @nesbitt2021 has been conducted to investigate the mechanisms for convective initiation and the occurrence of high-impact weather events associated with deep convection in central Argentina.

Forecasting mesoscale meteorological phenomena and particularly deep moist convection is a scientific and technological challenge due to its limited predictability, and the difficulties in diagnosing the state of the atmosphere at small spatial and short temporal scales (for example from 1 to 10 kilometers and on the order of minutes). Mesoscale data assimilation (DA) is an approach that can provide appropriate initial conditions for the initialization of high-resolution forecasts [@sun2014] and thus has received increasing attention in the last decades.

@@

In that regard, @wheatley2010 investigated the impact of assimilating surface pressure data in a mesoscale ensemble-based data assimilation system but found limited impact in two case studies related to mesoscale convective systems. @ha2014 showed that assimilating temperature and dew point temperature from high resolution surface weather station networks systematically improved the structure of the simulated planetary boundary layer and improved the short range precipitation forecast over the U.S. @chang2017, @bae2022 and @chen2016  reported beneficial effects of assimilating surface weather stations in a high resolution DA system using an ensemble Kalman filter, a 3-dimensional variational system and a 4-dimensional variational system respectively. Positive impacts have been found on the temperature and moisture forecast in the planetary boundary layer and in the location of precipitation systems. @sobash2015 showed in a mesoscale ensemble based DA system that the positive impact upon convection initiation and the short range precipitation forecast is achieved if data is assimilated frequently (in the order of minutes, rather than in the order of hours). @maejima2019 investigated the impact of 1-minute frequency assimilation of surface observations using observation simulation experiments on a heavy rainfall case founding that assimilating high frequency and spatially dense observations lead to a better representation of mesoscale circulation even though the number of observations provided by surface stations is much less than the one provided by weather radars. @gasperoni2018 conducted a case study to evaluate the impact of assimilating observations produced by private weather stations which are not incorporated in the operational analysis. They found a positive effect of these observations upon the initiation of deep moist convection along a dryline. This result is particularly important for data sparse regions such as Southern South America, where operational networks are not dense enough to capture mesoscale details.

The impact of other types of high spatial and temporal resolution observations, such as atmospheric motion vectors (AMVs) has been investigated in the context of limited-area mesoscale data assimilation. Many studies have focused on the impact of these observations on the prediction of tropical storms (e.g.  @wu2014, @cherubini2006, and @sawada2019, among many others). Most of these studies reported an overall positive impact of the assimilation of AMVs for this type of storm. However, some works indicated mixed impacts (e.g. @sawada2019 reported an improvement in the forecast of the track of the storm but a degradation in the forecasted intensity). As stated in @zhao2021 and @zhao2021a, the impact of assimilating these data on high impact weather events associated with mid-latitude deep convection over land has received relatively less attention. @zhao2021 and @zhao2021a assimilated GOES-16 AMVs into a storm-scale three-dimensional variational DA system during three high impact weather events. They reported positive impacts of AMVs on the characterization of the storm environment and improved short range precipitation forecasts. @otsuka2015 and @mallick2020 found a slight improvement in the short-range precipitation forecast due to the storm-scale assimilation of high frequency AMVs.

While the assimilation of radiance observations into global models is well established (), the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage, the bias correction, and the relatively low top models used for this application. @bao2015 studied the impact of assimilating microwave and infrared radiance data on temperature and humidity forecasts over the western USA and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data. More recently, @zhu2019 studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed an improvement for all variables, in particular for relative humidity at upper levels. @wang2021 studied the impact of assimilating radiances in the high-resolution Copernicus European Regional Reanalysis. They reported that satellite radiance observations had a neutral impact on the analyses of geopotential height in the lower troposphere, while a slightly negative impact on the upper troposphere and the stratosphere. They also observed similar results for 3-h forecasts initialized from the analysis but a positive impact on the short-range forecasts (12 and 24 h). Given the mixed results, there is still room to analyze the utility of assimilating radiance observations in a limited-area DA system over land. Moreover, to the best of the author's knowledge, there are few to none studies related to the direct assimilation of radiance observations over South America.

In particular, in South America, the relatively scarcity of conventional observations such as radiosondes increases the potential impact of the observation sources previously discussed. Previous work has shown promising results of mesoscale DA in South America using some of the above mentioned data sources [e.g. @dillon2016; @dillon2021; @goncalvesdegoncalves2015]. In particular @dillon2021 assimilated high resolution surface weather station networks, GOES-16 AMVs, and satellite temperature and moisture retrievals with promising results. However, no work has been conducted yet over this region to investigate the potential contribution of the datasets discussed above. 
The main objective of this work is to contribute to the quantification and comparison of the impact of high resolution surface weather stations, AMVs, and clear-sky satellite radiances, into a mesoscale, frequently-updated, and ensemble-based DA system. In particular, in this paper we will focus on the impact in the context of mid-latitude mesoscale convective system events. Another particular goal of this paper is to investigate the impact of these data sources in a region where the conventional observation network is rather sparse and where the potential contributions of these observing systems is larger.

@@

The paper is organized as follows. The data assimilation system, the experimental design, and the observations used are presented in section 2. Results are discussed in section 3 and finally, conclusions are summarized in section 4.

@@

On Nov 22, a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS. 
The MCS traveled approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 42 hours. 
To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MCS.

## Data assimilation system configuration

The model simulations for the case study are performed using version 3.9.1 of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW (V3.9), @skamarock2008) model. 
The horizontal grid spacing is 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the model at 50 hPa. 
The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25$^{\circ}$ horizontal grid spacing and 6-hour temporal resolution; @cisl_rda_ds084.1). 
The domain covers the area indicated in Figure \@ref(fig:dominio) to capture the development of the MCS during the simulated period.

The analyses are generated using the LETKF implementation (V1.3, @hunt2007) part of the Gridpoint Statistical Interpolation analysis system (GSI V3.8; @shao2016). 
A rapid update cycle approach is implemented with hourly analysis and a centered assimilation window, meaning that all the observations within $\pm$ 30 minutes of the analysis time are assimilated. 
Observations are assimilated in a 4D approach by comparing them with the corresponding first guess state at 10-minute intervals. 
For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; @han2006) is used as an observation operator to calculate model-simulated brightness temperatures. In addition to being already coupled to the GSI system.

We use a 60-member ensemble where the ensemble mean at the beginning of the DA cycle is initialized using the GFS deterministic analysis. Since the publicly available GFS ensemble members change at each cycle we used random perturbations to generate the initial ensemble perturbations. The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5$^{\circ}$ horizontal grid spacing with a smooth time evolution as in @necker2020 and @maldonado2021. In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales. This method helps to prevent an underestimation of the ensemble spread [@ouaraini2015], and the random perturbations are also applied at the boundaries to maintain proper levels of ensemble spread within the domain.

@@

(ref:dominio) a) The domain used for the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box) and the locations of Automatic Surface Weather Stations (AWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles). b) Locations of radiosonde launches during the RELAMPAGO. Green dots correspond to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8, and purple squares are radiosondes launched outside the IOP missions. The topography in meters is also shown (shaded).

@@

temp <- dominio %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = hgt), proj = norargentina_lambert,
                    breaks = seq(0, 6000, 500)) +
  scale_fill_gradient(low = "#f2f2f2", high = "#333333",
                      name = NULL,
                      breaks = seq(0, 6000, 1000),
                      guide = guide_colorstrip(barwidth = 20,
                                               barheight = 0.5)) +
  geom_mapa() +
  coord_sf(ylim = c(-35.5, -29), xlim = c(-66.5, -61.5)) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

@@

The conventional observations used are part of the Global Data Assimilation System (GDAS) data stream. We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at the National Centers for Environmental Prediction (NCEP). These consist of surface observations from 117 Conventional Surface Weather Stations (CSWS), ships, and upper-air observations from 13 radiosondes sites and aircraft. The orange triangles in Figure \@ref(fig:dominio) indicate the location of the surface station included in this experiment. The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for radiosondes. Wind surface observations over oceans (ASCATW) come from scatterometers and are also included in the PREPBUFR files.

@@

We also assimilate data from 866 Automatic Weather Stations (AWS) that are part of 17 public and private surface networks over Southern South America. The dataset used in this study has been obtained at the RELAMPAGO Data Set repository [@garcia2019]. These stations are indicated as green squares in Figure \@ref(fig:dominio)a. They have higher spatial coverage than the CSWS and a sampling frequency of 10 minutes in most cases. All stations measure temperature, but only 395 stations provide humidity, 422 provide pressure and 605 provide wind information. 
Observation errors used to assimilate these observations are the same as for the CSWS (see Table \@ref(tab:table-obs))

@@

Satellite-derived wind observations are also included in the PREPBUFR files available every 6 hs,  and consist of estimations from GOES-16 (using the visible, infrared, and water vapor channels) and METEOSAT 8 and 11 (using the visible and water vapor channels). Due to the domain covered by each of these satellites, GOES-16 is the primary source of satellite-derived winds (99 % of the observations). Observation errors used to assimilate these observations are indicated in Table \@ref(tab:table-obs).

```{r table-obs}
fread(here("analysis/data/derived_data/table1.csv")) %>% 
  # str()
  kbl(caption = "Characteristics of the assimilated observations: The code for each observation type and its source, the available variables, the observation error and the gross check thresholds used.",
      col.names = c("Code", "Platform", "Variable", "Error", "Gross check"),
      booktabs = TRUE,
      escape = FALSE) %>% 
  kable_classic_2(full_width = FALSE) %>% 
  kable_styling(font_size = 6,
                position = "center") %>% 
  column_spec(1, width = "3.5em") %>% 
  column_spec(2, width = "4.5em") %>% 
  column_spec(3, width = "5em") %>% 
  column_spec(4:5, width = "7em") %>% 
  collapse_rows(columns = 1:2, valign = "middle", latex_hline = "major") %>% 
  footnote(symbol = c("Observation error varied with height.",
                      "Observations above 600 hPa are rejected.",
                      "Observation error depends on the report type."),
           symbol_manual = c("*", "**", "+"))
```

@@

Satellite radiances available through the GDAS data stream, consisting of infrared and microwave observations are used in this study. This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Microwave Humidity Sounder (MHS), and 2 multispectral sensors; the Atmospheric Infrared Sounder (AIRS) and the Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table 2). Since the regional domain is located in the mid-latitudes and the satellite platforms of interest are on polar orbits, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath. For this reason, the number of satellite observations varied significantly among cycles. In particular, the multispectral sensors provided between 100 and 1000 observations for every scan every 12 hours, contributing 88 % of the total amount of assimilated radiances in our experiment. The vertical location of each radiance observation was estimated as the model level at which its weighting function was maximized as calculated by CRTM. The multispectral sensors have good vertical coverage and are able to sense from the lower troposphere and up to the lower stratosphere. 
The channels accepted for assimilation and their associated errors were defined taking into account the low model top (50 hPa). The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically. Firstly a spatial data thinning is applied using a 60 km grid following [@singh2016; @jones2013; @lin2017a], where the observations to be assimilated are chosen based on their distance to the model grid points, the observation quality (based on available data quality information), and the number of available channels (from the same pixel and sensor) that passed the quality control. Also, observations over the sea are preferred to those over land or snow [@hu2018].

The thinned observations were then bias corrected. The bias correction (BC) has an air-mass dependent and an angle-dependent component [@zhu2014] and it is calculated as a multi-linear function of N predictors $p_i(x)$, with associated coefficients $\beta_i$. Then, the bias corrected brightness temperature (BT_{bc}) can be obtained as:

@@

In the GSI system, the $\beta_i$ coefficients are trained using a variational estimation method which solves the $\beta_i$ that provides the best fit between the simulation and the observations. The coefficients were initialized at 18 UTC Nov 18 with the GFS system coefficients. The assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments in the estimated coefficients at each time.

@@

(ref:obs-cycle) a) Number of assimilated observations per cycle and b) time averaged number of assimilated observations per cycle divided into 50 hPa-depth vertical layers for the CONV (blue squares and line), AWS (light blue dots and line), SATWND (orange triangles and line) and RAD (red diamond and line) experiments.

@@

where $O$ and $X$ stand for independent observations and the simulations respectively.

>>>>

---
title: "Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a  Mesoscale Convective System Case During RELAMPAGO campaign"
author:
- name: Paola Belén Corrales
  email: paola.corrales@cima.fcen.uba.ar
  affiliation: UBA,CIMA,CNRS
  footnote: 1
- name: Victoria Galligani
  affiliation: UBA,CIMA,CNRS
- name: Juan Ruiz
  affiliation: UBA,CIMA,CNRS
- name: Luiz Sapucci
  affiliation: INPE
- name: María Eugenia Dillon
  affiliation: SMN,CONICET
- name: Yanina García Skabar
  affiliation: SMN,CONICET,CNRS
- name: Maximiliano Sacco
  affiliation: SMN
- name: Craig S. Schwartz
  affiliation: NCAR
- name: Stephen W. Nesbitt
  affiliation: Illinois
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_book: 
      base_format: rticles::elsevier_article
abstract: |
  In this paper, we evaluate the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF over central and north eastern Argentina where the surface and upper air observing networks are relatively coarse. We conducted a case study corresponding to a huge mesoscale convective system (MCS) that developed during November 22, 2018. The accumulated precipitation associated with this MCS was quite high, exceeding 200 mm over northern Argentina and Paraguay. The MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign.

  We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-member multiphysics ensemble. We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files; AWS combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds and RAD, including SATWND, and satellite radiances from different microwave and infrared sensors. We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study. The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation. We also run ensemble forecasts initialized from each experiment and evaluated their skill to predict precipitation. We found that the hourly assimilation of the observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.
journal: Atmospheric Research
#layout: authoryear,preprint,review,12pt
layout: final,5p,times,twocolumn,authoryear
address:
- code: UBA
  address: Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmósfera y los Océanos. Buenos Aires, Argentina.
- code: CIMA
  address: CONICET – Universidad de Buenos Aires. Centro de Investigaciones del Mar y la Atmósfera (CIMA). Buenos Aires, Argentina.
- code: CNRS
  address: CNRS – IRD – CONICET – UBA. Instituto Franco-Argentino para el Estudio del Clima y sus Impactos (IRL 3351 IFAECI). Buenos Aires, Argentina.
- code: SMN
  address: Servicio Meteorológico Nacional de Argentina.
- code: CONICET
  address: CONICET (Consejo Nacional de Investigaciones Científicas y Técnicas).
- code: INPE
  address: National Institute for Space Research, Brazil, Center for Weather Forecasting and Climate Studies.
- code: NCAR
  address: National Center for Atmospheric Research, Boulder, Colorado.
- code: Illinois
  address: Department of Atmospheric Sciences, University of Illinois at Urbana–Champaign, Urbana, Illinois.
bibliography: 
  - paper_ANA.bib 
  - packages.bib
csl: elsevier-harvard.csl
footnote:
- code: 1
  text: Corresponding Author
keywords: Regional Data Assimilation, Surface Obsevations, Satellite Observations 
numbersections: true
# header-includes: 
#  - \usepackage{gensymb}
#  - \usepackage{subfig}
#  - \usepackage[inline]{showlabels}
#  - \usepackage{chngcntr}
#  - \usepackage{natbib}
#  - \usepackage{lineno}
#  - \linenumbers 
---

@@

Severe weather events cause significant human and material losses around the world. A large number of these phenomena are associated with the occurrence of deep moist convection including tornadoes, intense wind gusts, extreme precipitation in short time periods, large hail, and lightning. 
Southern South America has one of the highest frequencies in the world of favorable conditions for high-impact meteorological events [@brooks2003] and large hail events [@cecil2012], particularly during austral spring and summer. 
This is also confirmed by observational evidence and high impact weather reports [@matsudo2015; @rasmussen2014]. Recently, the RELAMPAGO-CACTI field campaign [@nesbitt2021] has been conducted to investigate the mechanisms for convective initiation and the occurrence of high-impact weather events associated with deep convection in central Argentina.

Forecasting mesoscale meteorological phenomena and particularly deep moist convection is a scientific and technological challenge due to its limited predictability and the difficulties in diagnosing the state of the atmosphere at small spatial and short temporal scales (for example from 1 to 10 kilometers and on the order of minutes). Mesoscale data assimilation (DA) is an approach that can provide appropriate initial conditions for high-resolution numerical forecasts [@sun2014] and thus has received increasing attention in the last decades.

@@

In that regard, several authors investigated the impact of assimilating surface weather data (e.g. @wheatley2010, @ha2014, @chang2017, @bae2022, @banos2021, @maejima2019, and @chen2016 using different assimilation methodologies. Most of these studies reported the beneficial impacts of assimilating temperature and dew point observations upon the planetary boundary layer structure and the location and timing of precipitating systems.  @sobash2015 showed using a mesoscale DA system, that the positive impact upon convection initiation and the short range precipitation forecast is achieved if data is assimilated frequently (in the order of minutes, rather than in the order of hours).  More recently, @gasperoni2018 evaluated the impact of assimilating observations produced by private weather station networks which are not incorporated in the operational analysis. They found a positive effect of these observations upon the initiation of deep moist convection. This result is particularly important for data sparse regions such as Southern South America, where operational networks are not dense enough to capture mesoscale details.

The impact of other types of high spatial and temporal resolution observations, such as atmospheric motion vectors (AMVs), has also been investigated in the context of limited-area mesoscale DA. Many studies have focused on the impact of these observations on the prediction of tropical storms (e.g.,  @wu2014, @cherubini2006, and @sawada2019, among many others). Most of these studies reported an overall positive impact of the assimilation of AMVs for this type of storm. However, some works indicated mixed impacts (e.g. @sawada2019 reported an improvement in the forecast of the track of the storm but a degradation in the forecast intensity). As stated in @zhao2021 and @zhao2021a, the impact of assimilating these data on high impact weather events associated with mid-latitude deep convection over land has received relatively less attention. @zhao2021 and @zhao2021a assimilated GOES-16 AMVs into a storm-scale three-dimensional variational DA system during three high impact weather events. They reported positive impacts of AMVs on the characterization of the storm environment and improved short range precipitation forecasts. @otsuka2015 and @mallick2020 found a slight improvement in the short-range precipitation forecast due to the storm-scale assimilation of high frequency AMVs.

While the assimilation of radiance observations into global models is well established [@eyre2020], the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage, the bias correction, and the relatively low model tops used for this application. @bao2015 studied the impact of assimilating microwave and infrared radiance data on temperature and humidity forecasts over the western USA and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data. More recently, @zhu2019 studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed an improvement for all variables, in particular for relative humidity at upper levels. @wang2021 studied the impact of assimilating radiances in the high-resolution Copernicus European Regional Reanalysis. They reported that satellite radiance observations had a neutral impact on the analyses of geopotential height in the lower troposphere, while a slightly negative impact on the upper troposphere and the stratosphere. They also observed similar results for 3-h forecasts initialized from the analysis but a positive impact on 12 and 24 -h forecasts. Given the mixed results, there is still room to analyze the utility of assimilating radiance observations in a limited-area DA system over land. Moreover, to the best of our knowledge, there are none studies related to the direct assimilation of radiance observations over South America.

In particular, in South America, the relatively scarcity of conventional observations such as radiosondes increases the potential impact of the observation sources previously discussed. Previous work has shown promising results of mesoscale DA in South America using some of the above mentioned data sources [e.g. @dillon2016; @dillon2021; @goncalvesdegoncalves2015]. In particular @dillon2021 assimilated high resolution surface weather station networks, GOES-16 AMVs, and satellite temperature and moisture retrievals with promising results. However, no work has been conducted yet over this region to investigate the potential contribution of the datasets discussed above. 
The main objective of this work is to contribute to the quantification and comparison of the impact of high resolution surface weather stations, AMVs, and clear-sky satellite radiances, into a mesoscale, frequently-updated ensemble-based DA system. In particular, we will focus on the impact in the context of mid-latitude mesoscale convective system events. Another particular goal of this paper is to investigate the impact of these data sources in a region where the conventional observation network is rather sparse and where the potential contributions of these observing systems is larger.

@@

The paper is organized as follows. The DA system, the experimental design, and the observations used are presented in section 2. Results are discussed in section 3 and finally, conclusions are summarized in section 4.

@@

On Nov 22, 2018 a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS. To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MCS. 
The MCS traveled approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 42 hours.

## Data assimilation system configuration {#config}

The model simulations for the case study are performed using version 3.9.1 of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW, @skamarock2008) model. 
The horizontal grid spacing is 10 km (150 x 200 grid points) in the horizontal and 37 levels in the vertical with the top of the model at 50 hPa. 
The initial and boundary conditions are provided by the Global Forecast System (GFS) analysis (0.25$^{\circ}$ horizontal grid spacing and 6-hour temporal resolution; @cisl_rda_ds084.1). 
The domain covers the area indicated in Figure \@ref(fig:dominio) to capture the development of the MCS during the simulated period.

The analyses are generated using the LETKF implementation (V1.3, @hunt2007) part of the Gridpoint Statistical Interpolation analysis system (GSI V3.8; @shao2016). 
A rapid update cycle approach is implemented with hourly analysis and a centered assimilation window, meaning that all the observations within $\pm$ 30 minutes of the analysis time are assimilated. 
Observations are assimilated in a 4D approach by comparing them with the corresponding first guess state at 10-minute intervals. 
For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; @han2006) is used as an observation operator to calculate model-simulated brightness temperatures.

We use a 60-member ensemble where the ensemble mean at the beginning of the DA cycle is initialized using the GFS deterministic analysis. with random perturbations to generate the initial ensemble perturbations. The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5$^{\circ}$ horizontal grid spacing with a smooth time evolution as in @necker2020 and @maldonado2021. In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales. This method helps to prevent an underestimation of the ensemble spread [@ouaraini2015], and the random perturbations are also applied at the boundaries to maintain proper levels of ensemble spread within the domain.

@@

(ref:dominio) a) The domain used for the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box), and the locations of Automatic Weather Stations (AWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles). b) Locations of radiosonde launches during the RELAMPAGO. Green dots correspond to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8, and purple squares are radiosondes launched outside the IOP missions. The topography in meters is also shown (shaded).

@@

temp <- dominio %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = hgt), proj = norargentina_lambert,
                    breaks = seq(0, 6000, 500)) +
  scale_fill_gradient(low = "#f2f2f2", high = "#333333",
                      name = "Altitude (m)",
                      breaks = seq(0, 6000, 1000),
                      guide = guide_colorstrip(barwidth = 20,
                                               barheight = 0.5)) +
  geom_mapa() +
  coord_sf(ylim = c(-35.5, -29), xlim = c(-66.5, -61.5)) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

@@

The conventional observations used are part of the Global Data Assimilation System (GDAS) data stream. We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at the National Centers for Environmental Prediction (NCEP). These consist of surface observations from 117 Conventional Surface Weather Stations (CSWS), ships, and upper-air observations from 13 radiosondes sites and aircraft. The orange triangles in Figure \@ref(fig:dominio) indicate the location of the surface stations included in this experiment. The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for radiosondes. Wind surface observations over oceans (ASCATW) come from scatterometers and are also included in the PREPBUFR files.

@@

We also assimilate data from 866 Automatic Weather Stations (AWS) that are part of 17 public and private surface networks over Southern South America. The dataset used in this study has been obtained from the RELAMPAGO Data Set repository [@garcia2019]. These stations are indicated as green squares in Figure \@ref(fig:dominio)a. They have higher spatial coverage than the CSWS and a sampling frequency of 10 minutes in most cases. All stations measure temperature, but only 395 stations provide humidity, 422 provide pressure, and 605 provide wind information. 
Observation errors used to assimilate these observations are the same as for the CSWS (see Table \@ref(tab:table-obs)).

@@

Satellite-derived wind observations are also included in the PREPBUFR files available every 6 h,  and consist of estimations from GOES-16 (using the visible, infrared, and water vapor channels) and METEOSAT 8 and 11 (using the visible and water vapor channels). Due to the domain covered by each of these satellites, GOES-16 is the primary source of satellite-derived winds (99 % of the observations). Observation errors used to assimilate these observations are indicated in Table \@ref(tab:table-obs).

```{r table-obs}
fread(here("analysis/data/derived_data/table1.csv")) %>% 
  # str()
  kbl(caption = "Characteristics of the assimilated observations: The code for each observation type and its source, the available variables, the observation error, and the gross check thresholds used.",
      col.names = c("Code", "Platform", "Variable", "Error", "Gross check"),
      booktabs = TRUE,
      escape = FALSE) %>% 
  kable_classic_2(full_width = FALSE) %>% 
  kable_styling(font_size = 6,
                position = "center") %>% 
  column_spec(1, width = "3.5em") %>% 
  column_spec(2, width = "4.5em") %>% 
  column_spec(3, width = "5em") %>% 
  column_spec(4:5, width = "7em") %>% 
  collapse_rows(columns = 1:2, valign = "middle", latex_hline = "major") %>% 
  footnote(symbol = c("Observation error varied with height.",
                      "Observations above 600 hPa are rejected.",
                      "Observation error depends on the report type."),
           symbol_manual = c("*", "**", "+"))
```

@@

Satellite radiances available through the GDAS data stream, consisting of infrared and microwave observations, are used in this study. This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Microwave Humidity Sounder (MHS), and 2 multispectral sensors; the Atmospheric Infrared Sounder (AIRS) and the Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table 2). Since the regional domain is located in the mid-latitudes and the satellite platforms of interest are on polar orbits, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath. For this reason, the number of satellite observations varied significantly among cycles. In particular, the multispectral sensors provided between 100 and 1000 observations for every scan every 12 hours, contributing 88 % of the total amount of assimilated radiances in our experiment. The vertical location of each radiance observation was estimated as the model level at which its weighting function was maximized as calculated by CRTM. The multispectral sensors have good vertical coverage and are able to sense from the lower troposphere up to the lower stratosphere. 
The channels accepted for assimilation and their associated errors were defined taking into account the low model top (50 hPa). The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically. First, a spatial data thinning is applied using a 60 km grid following [@singh2016; @jones2013; @lin2017a], where the observations to be assimilated are chosen based on their distance to the model grid points, the observation quality (based on available data quality information), and the number of available channels (from the same pixel and sensor) that passed the quality control. Also, observations over the sea are preferred to those over land or snow [@hu2018].

The thinned observations were then bias corrected. The bias correction (BC) has an air-mass dependent and an angle-dependent component [@zhu2014] and it is calculated as a multi-linear function of N predictors $p_i(x)$, with associated coefficients $\beta_i$. Then, the bias corrected brightness temperature ($BT_{bc}$) can be obtained as:

@@

In the GSI system, the $\beta_i$ coefficients are trained using a variational estimation method which solves for the $\beta_i$ that provides the best fit between the simulation and the observations. The coefficients were initialized at 18 UTC Nov 18, 2018 with the GFS system coefficients. The assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments in the estimated coefficients at each time.

@@

(ref:obs-cycle) a) Number of assimilated observations per cycle and b) time averaged number of assimilated observations per cycle divided into 50 hPa-depth vertical layers for the CONV (blue squares and line), AWS (light blue dots and line), SATWND (orange triangles and line) and RAD (red diamonds and line) experiments.

@@

where $O$ and $X$ stand for independent observations, and the simulations respectively and N is the sample size.
